Species sympatry shapes brain size evolution in Primates

### 2021-12-14 Article/Coverletter.Rmd
Update article + update readme + reorganise files

|   The road to understanding the evolutionary history of cognition is still little paved. Currently formulated theories generally oppose social and ecological drivers of cognition and are often framed within a foraging-related context. These theories, nonetheless, primarily focus on the challenges an individual faces with regards to its food spatio-temporal distribution, or to its conspecifics from a same group and/or species. Space is also shared with other species with similar ecological niche that might also shape the environment and be competitors or cooperators. Thus, it is likely that cognate species that share diet are likely to affect the social and ecological environment a species evolves in. This should therefore loop onto the selective pressures that apply to cognition. This facet remained unexplored.

|   To help filling this gap, we perform a global-scale analysis to investigate the role of sympatry in shaping the size of different brain areas in animals, and shaping their evolutionary success. In particular, we focus on frugivorous primates because the consumption of fruits is highly tightened to advanced cognitive skills. Using recently developed phylogenetic models accounting for competition or not with sympatric species, we retrace the evolution of the size of brain areas related to foraging or not. We also correlate brain size, diversification rate (a proxy for evolutionary success), and sympatry intensity.

|  We show that the evolutionary history of brain areas associated to foraging tasks is more accurately described by competitive models. In particular, sympatry negatively affects their relative size. We additionally show that, even if these brain sizes are not related to diversification rate, this latter is reduced when sympatry rates increases. Altogether, these findings suggest that sympatry is associated to a relaxed selection for bigger brain, or even a positive selection for smaller brain. We speculate that this is due to an over-complexification of the environment, which makes cognition not advantageous anymore.

|  This study enlarges our understanding on the effect of space use and species distribution in animal cognition. It shows unprecedent evidence of direct and/or indirect effect of competition on cognition evolution, questioning proximate mechanisms at play. We strongly think that this stands as an important complementary piece of global analyses that were already published in your journal that will be of interest to empiricists and theoreticians from ecological and evolutionary fields. On behalf of all authors, I certify that this is a completely novel contribution that has not been submitted elsewhere, and that all authors approved the manuscript and its submission to *`r journal`*.

>>>>

|   The road to understanding the evolutionary history of cognition is still little paved. Currently formulated theories generally oppose social and ecological drivers of cognition, and prioritise direct or indirect interactions with conspecifics. Yet, space is also shared with other species with similar ecological niches. Thus, it is likely that cognate species that compete for food are likely to affect the social and ecological environment a species evolves in too. This should therefore loop onto the selective pressures that apply to cognition. This facet remained unexplored.

|   To help fill this gap, we perform a global-scale analysis to investigate the role of sympatry in shaping the size of different brain areas, and shaping their evolutionary success. In particular, we focus on frugivorous primates because the consumption of fruits is highly tightened to advanced cognitive skills. Using recently developed phylogenetic models accounting for species sympatry or not, we retrace the evolution of the size of brain areas related to foraging or not. We also correlate brain sizes, species diversification rates, and levels of sympatry.

|  We show that the evolutionary history of brain areas associated with foraging tasks is more accurately described by models considering  sympatry. In particular, sympatry negatively affects their relative size. We additionally show that, even if these brain sizes do not directly impact primate diversification, this latter slowdowns when sympatry increases. Altogether, these findings suggest that sympatry is associated with a positive selection for smaller brains. We speculate that this is due to an over-complexification of the environment, which makes cognition not advantageous anymore for sympatric species.

|  This study enlarges our understanding of the effect of space use and species distribution in animal cognition. It shows unprecedented evidence of the direct and/or indirect effects of interspecific competition on cognition evolution, questioning proximate mechanisms at play. We strongly think that this stands as an important complementary piece of global analyses that were already published in your journal. We believe it will be of interest to empiricists and theoreticians from ecological and evolutionary fields. On behalf of all authors, I certify that this is a completely novel contribution that has not been submitted elsewhere, and that all authors approved the manuscript and its submission to *`r journal`*.
